When are Fuzzy Extractors Possible?

### 2014-09-15 trunk/info-bounds/FuzzyExtPossibility.tex
writing down distribution aware impossibility

\bnote{connection to obfuscation}
\section{Preliminaries}
\label{sec:preliminaries}
\bnote{be careful with overloading of $\ell$, both key length and number of levels.}
For a random variables $X_i$ over some alphabet $\mathcal{Z}$ we denote by $X = X_1,..., X_\ell$  the tuple $(X_1,\dots, X_\ell)$.  For a set of indices $J$, $X_{J}$ is the restriction of $X$ to the indices in $J$.  The set $J^c$ is the complement of $J$.  The {\em min-entropy} of $X$ is $\Hoo(X) = -\log(\max_x \Pr[X=x])$,
and the {\em average (conditional)} min-entropy of $X$ given $Y$ is  $\Hav(X|Y) = -\log(\expe_{y\in Y} \max_{x} \Pr[X=x|Y=y])$~\cite[Section 2.4]{DBLP:journals/siamcomp/DodisORS08}.   For a random variable $W$, let $H_0(W)$ be the logarithm of the size of the support of $W$,  that is $H_0(W) = \log |\{w | \Pr[W=w]>0\}|$.  We will also use an average case version of this notion~(to the best of our understanding this has not been used before).\bnote{has it?}
\begin{definition}[Conditional Hartley Entropy]
\label{def:conditional hartley}
The conditional Hartley entropy of $X |Y $ is $\tilde{H}_0(X |Y) = \log ( \expe_{y\in Y} 2^{-H_0(X | Y=y)})$.
\end{definition}
The {\em statistical distance} between random variables $X$ and $Y$ with the same domain is $\Delta(X,Y) = \frac12 \sum_x |\Pr[X=x] - \Pr[Y=x]|$.
For a distinguisher $D$ we write the \emph{computational distance} between $X$ and $Y$ as $\delta^D(X,Y) = \left| \expe[D(X)]-\expe[D(Y)]\right |$ (we extend it to a class of distinguishers $\mathcal{D}$ by taking the maximum over all distinguishers $D\in\mathcal{D}$).  We denote by $\mathcal{D}_{s}$ the class of randomized circuits which output a single bit and have size at most $s$.

@@

\section{Distribution-Dependent Sketches}
In this section, we consider show it is possible to build distribution-aware fuzzy extractors and secure sketches whenever $\Hfuzz(W)= \omega(\log n)$.
We consider increasingly complex types of distributions and end with an arbitrary distribution over a metric space.  We construct secure sketches but our constructions can be transformed to fuzzy extractors using \lemref{lem:fuzzy ext construction}.  In \secref{sec:dist oblivious} we consider distribution-oblivious algorithms.

@@

\begin{theorem}
Let $n$ be a security parameter.  There exists a family of distributions $\mathcal{V}$ such that for each element $W\in \mathcal{V}$, $\Hfuzz(W)= \omega(\log n)$ and there is no $(\mathcal{M}, \mathcal{V}, \tilde{m}, t)$-secure sketch with error $\delta \le 1/2$ for $\tilde{m} =\omega(1)$.  %Furthermore, consider $V$ as the uniform distribution over $\mathcal{V}$.  There is no secure sketch for $\mathcal{
\bnote{what is $t$?}\bnote{finish proof}
\end{theorem}

\begin{proof}
\bnote{need to rearrange this proof.}
It will be convenient to think of a single distribution where the outcome $w$ is a two step process of first sampling $W\leftarrow \mathcal{W}$ and then sampling $w\leftarrow W$.  The adversary sees the information $W$, but the system designer does not.  We can think of the fuzzy extractor working for a distribution $V$ and the adversary receiving some side information $Z$~(which distribution was sampled).

@@

Let $\mathcal{V}$ be family of distributions of the form $W = w_1,..., w_\ell$ where $w_1$ is uniform and $w_i = a_i w_1 + b_i$ for $2\le i \le \ell$ and $a_i\neq 0$.  This type of distribution is a line in space $\mathbb{F}^\ell$.  Define $V$ as the uniform distribution over all such $W$.  Fix some $1< t < \ell$.
It suffices to show the following:

\begin{itemize}
\item For all $W\in \mathcal{V}$, $\Hfuzz(W) = \omega (\log n)$.
\item The distribution $V$ is uniform~(from the sketch designers point of view).  That is $V\overset{d}=U$.
\item For any secure sketch on $V$, $H_0(V | \sketch(V)) \le n -\log |B_t|$. 
\item The information $Z$ is a universal hash of $V$ that is independent of $\sketch(V)$. This suffices to show that $\Hav(V| \sketch(V), Z)$ is very small.
\end{itemize}
\begin{proposition} For each $W\in\mathcal{V}$, $\Hfuzz(W) = \omega(\log n)$.
\end{proposition}
\begin{proof}
Consider some fixed $W\in\mathcal{V}$.  The value $w_1$ is uniform in a field of size $\omega(\poly(n))$, so $\Hoo(W) =\omega(\log n)$.  Recall that $t< \ell$.  We now show that for any $w, w'\in W$, $\dis(w, w') = \ell$ and thus $\Hfuzz(W) = \Hoo(W)$.  Fix some $w, w'\in W$.  Clearly, $w_1 \neq w_1'$, for any $i$, $w_i = a_i w_1 + b_i$ and $w_i' = a_i w_1' + b_i$.  Since $a_i\neq 0$, $a_iw_1 \neq a_iw_1'$ and thus $a_iw_1+b_i \neq a_iw_1'+b_i$.  That is, $\dis (w, w')  =\ell$.
\end{proof}

\begin{proposition}$V$ is the uniform distribution over $\mathbb{F}^\ell$.
\end{proposition}
\begin{proof}
Consider some $w\in V$.  Then $w$ was drawn from some intermediate distribution $W$ with coefficients $a_2, b_2, ..., a_\ell , b_\ell$.  The value $w_1$ is uniformly random and $w_i$ are uniformly random since $b_2,..., b_\ell$ are uniformly random.
\end{proof}
\bnote{will need to define average Hartley entropy}
\bnote{show version without error first?}
\begin{proposition}
Fix some $\sketch, \rec$ algorithm with error $\delta < 1/4$, then $H_0(V | \sketch(V)) \le (\ell-t) |\mathbb{F}|n +1$.\footnote{This result is an extension of lower bounds from~\cite[Appendix C]{DBLP:journals/siamcomp/DodisORS08}.  Dealing with imperfect correctness makes the bound more complicated.  In particular, we can only argue about the average remaining support size.}
\end{proposition}
\begin{proof}
We assume that $\rec$ is deterministic in our analysis.  Any randomness necessary for the \rec algorithm can be provided by \sketch.  We will however, not consider these coins in our security analysis.

@@

Consider some $p\in A_{ss}$.  We now show that $H_0(V | p) \le n-\log |B_t|$.  Note for the true value $w$ sketched $\{w' | \dis(w, w') \le t \wedge \rec(w', p) \neq w] \le 2\delta$.  Thus for every value in $V|p$ this is also true.  This makes the support of $V|p$ a $(t, 2\delta)$-Shannon code~(see \defref{def:shannon-code}).  In particular, this implies that for all $w_1, w_2 \in V|p$, $\dis(w_1, w_2)\ge t$~(since $2\delta< 1/2$).  That is $V|p$ is a set with minimum distance at least $t$.  By the Singleton bound, this implies that $H_0(V | p) \le (\ell -t )|\mathbb{F}|$.  Averaging over $p$ one has that $\tilde{H}_0(V|P) \le (\ell -t ) |\mathbb{F}| +1$.
%By the Markov inequality, 
%\[
%\forall w, w^*, \Pr_{p \leftarrow \sketch(w)} [ \Pr[\rec(w^*, p) = w] \le 1-2\delta ] < \delta /2.
%\]
%That is, for most $p$, $\Pr[\rec(w^*, p) = w] > 1-2\delta>1/2$.
%
%Outline of what I want to say.
%\begin{itemize}
%\item Suppose we are at some good sketch value, that is, the sketch decodes well for $w^*$ close to the true value $w$.
%\item Then for the true value that was sketched, for most neighbors $w^*$ of $w$, $\Pr[\rec(w^*, p) =w] >1/2$.  
%\item For a value $w'$ to be in the support of $W | P=p$, it must be for all neighbors of $\Pr[\rec( w', p) =w ]>1/2$.
%\item This implies that if $w_1, w_2 \in W|P = p$, then $\dis (w_1, w_2)\ge t$.
%\item Conditioned on $P=p$, $W| P=p$ is a good code for random errors.
%Consider some subset $G\subseteq supp(W| P=p)$ with $\Pr (W| P=p \in G)\ge 1/2$.  Then for all $w\in G$ the probability they recover correctly is at least $1-4\delta$~(by another application of Markov).  
%\item This, means that the size of $|G|$ is at most $n -\log |B_t|$.
%\end{itemize}
\end{proof}
We now define the side information $Z = a_2, b_2, ..., a_\ell, b_\ell$.  By described above the view of $\sketch, \rec$ is a uniform distribution $V$.  Thus, the side information $Z$ is independent of $\sketch(V)$.
\begin{lemma}
$\Hav(V| P , Z) = \Theta(1)$
\end{lemma}
\begin{proof}
Consider some fixed $p$.
If $Z$ is universal then $\forall w_1, w_2 \in V|P=p, w_1\neq w_2$\bnote{why should $Z$ be universal for the entire range, I get for each coordinate.}
\[
\Pr[Z(w_1) = Z(w_2)] = \frac{1}{|Z|}.
\]
Then by the union bound, 
\[
\Pr[\exists w_1\neq w_2 \wedge Z(w_1) = Z(w_2)]\le \frac{ 2^{H_0(V|P=p)}}{|Z|}.
\]
\bnote{need to average across $P$}
\bnote{I'm confused about what we mean by $Z$ being universal here.}
\bnote{working here!}
\end{proof}
%
%\bnote{stopped here}
\end{proof}

>>>>

\bnote{connection to obfuscation}
\section{Preliminaries}
\label{sec:preliminaries}
\bnote{be careful with overloading of $\ell$, both key length and number of levels.}
For a random variables $X_i$ over some alphabet $\mathcal{Z}$ we denote by $X = X_1,..., X_\ell$  the tuple $(X_1,\dots, X_\ell)$.  For a set of indices $J$, $X_{J}$ is the restriction of $X$ to the indices in $J$.  The set $J^c$ is the complement of $J$.  The {\em min-entropy} of $X$ is $\Hoo(X) = -\log(\max_x \Pr[X=x])$,
and the {\em average (conditional)} min-entropy of $X$ given $Y$ is  $\Hav(X|Y) = -\log(\expe_{y\in Y} \max_{x} \Pr[X=x|Y=y])$~\cite[Section 2.4]{DBLP:journals/siamcomp/DodisORS08}.   For a random variable $W$, let $H_0(W)$ be the logarithm of the size of the support of $W$,  that is $H_0(W) = \log |\{w | \Pr[W=w]>0\}|$.  We will also use an average case version of this notion~(to the best of our understanding this has not been used before):\bnote{has it?}
\begin{definition}[Conditional Hartley Entropy]
\label{def:conditional hartley}
The conditional Hartley entropy of $X |Y $ is $\tilde{H}_0(X |Y) = \log ( \expe_{y\in Y} |\{x | \Pr[X=x |Y=y]>0\}|)$.
\end{definition}
The {\em statistical distance} between random variables $X$ and $Y$ with the same domain is $\Delta(X,Y) = \frac12 \sum_x |\Pr[X=x] - \Pr[Y=x]|$.
For a distinguisher $D$ we write the \emph{computational distance} between $X$ and $Y$ as $\delta^D(X,Y) = \left| \expe[D(X)]-\expe[D(Y)]\right |$ (we extend it to a class of distinguishers $\mathcal{D}$ by taking the maximum over all distinguishers $D\in\mathcal{D}$).  We denote by $\mathcal{D}_{s}$ the class of randomized circuits which output a single bit and have size at most $s$.

@@

\section{Distribution-Aware  Sketches}
In this section, we consider show it is possible to build distribution-aware fuzzy extractors and secure sketches whenever $\Hfuzz(W)= \omega(\log n)$.
We consider increasingly complex types of distributions and end with an arbitrary distribution over a metric space.  We construct secure sketches but our constructions can be transformed to fuzzy extractors using \lemref{lem:fuzzy ext construction}.  In \secref{sec:dist oblivious} we consider distribution-oblivious algorithms.

@@

\begin{theorem}
Let $n$ be a security parameter.  There exists a family of distributions $\mathcal{V}$ such that for each element $W\in \mathcal{V}$, $\Hfuzz(W)= \omega(\log n)$ and there is no $(\mathcal{M}, \mathcal{V}, \tilde{m}, t)$-secure sketch with error $\delta < 1/4$ for $\tilde{m} =\omega(1)$ for $t\ge 4$.  \bnote{finish proof}
\end{theorem}

\begin{proof}
It will be convenient to think of a single distribution where the outcome $w$ is a two step process of first sampling $W\leftarrow \mathcal{W}$ and then sampling $w\leftarrow W$.  The adversary sees the information $W$, but the system designer does not.  We can think of the fuzzy extractor working for a distribution $V$ and the adversary receiving some side information $Z$~(which distribution was sampled).

@@

Let $\mathcal{V}$ be family of distributions of the form $W = w_1,..., w_\ell$ where $w_1$ is uniform and $w_i = a_i w_1 + b_i$ for $2\le i \le \ell$ and $a_i\neq 0$.  This type of distribution is an affine line in space $\mathbb{F}^\ell$.  Define $V$ as the uniform distribution over all such $W$.  Furthermore, the side information $Z = a_2, b_2, ..., a_\ell, b_\ell$. Fix some $4\le t < \ell$.
We then show the following:

\begin{itemize}
\item For all $W\in \mathcal{V}$, $\Hfuzz(W) = \omega (\log n)$. Alternatively, $\forall z, \Hfuzz(V | Z=z) = \omega(\log n)$.
\item The distribution $V$ is uniform~(from the sketch designers point of view).  That is $V\overset{d}=U$.
\item For any secure sketch on $V$, $\tilde{H}_0(V | \sketch(V)) \le  (\ell-t+1)\log |\mathbb{F}|$. 
\item Having seen $v\in V$, there are many possible values for $z_1, z_2$ for the side information $Z$~(and these possible values are equally likely).  Furthermore, the distributions $V|Z=z_1 $ and $V| Z=z_1$ have disjoint support outside of $v$.
\item For most values of possible $Z$ there are few values possible in the distribution $V | \sketch(V)$.  Thus, $\Hav(V| \sketch(V) , Z)$ is small.
\end{itemize}
\begin{proposition} 
\label{prop:each element good} For each $W\in\mathcal{V}$, $\Hfuzz(W) = \omega(\log n)$.
\end{proposition}
\begin{proof}
Consider some fixed $W\in\mathcal{V}$.  The value $w_1$ is uniform in a field of size $\omega(\poly(n))$, so $\Hoo(W) =\omega(\log n)$.  Recall that $t< \ell$.  We now show that for any $w, w'\in W$, $\dis(w, w') = \ell$ and thus $\Hfuzz(W) = \Hoo(W)$.  Fix some $w, w'\in W$.  Clearly, $w_1 \neq w_1'$, for any $i$, $w_i = a_i w_1 + b_i$ and $w_i' = a_i w_1' + b_i$.  Since $a_i\neq 0$, $a_iw_1 \neq a_iw_1'$ and thus $a_iw_1+b_i \neq a_iw_1'+b_i$.  That is, $\dis (w, w')  =\ell$.
\end{proof}

\begin{proposition}$V$ is the uniform distribution over $\mathbb{F}^\ell$.
\end{proposition}
\begin{proof}
Consider some $w\in V$.  Then $w$ was drawn from some intermediate distribution $W$ with coefficients $a_2, b_2, ..., a_\ell , b_\ell$.  The value $w_1$ is uniformly random and $w_i$ are uniformly random since $b_2,..., b_\ell$ are uniformly random.
\end{proof}
\bnote{show version without error first?}
\begin{proposition}
Fix some $\sketch, \rec$ algorithm with error $\delta < 1/4$, then $\tilde{H}_0(V | \sketch(V)) \le (\ell-t+1)\log |\mathbb{F}|+1$.\footnote{This result is an extension of lower bounds from~\cite[Appendix C]{DBLP:journals/siamcomp/DodisORS08}.  Dealing with imperfect correctness makes the bound more complicated.  In particular, we can only argue about the average remaining support size.}
\end{proposition}
\begin{proof}
We assume that $\rec$ is deterministic in our analysis.  Any randomness necessary for the \rec algorithm can be provided by \sketch.  We will however, not consider these coins in our security analysis.

@@

Consider some $p\in A_{ss}$.  We now show that $H_0(V | p) \le n-\log |B_t|$.  Note for the true value $w$ sketched $\{w' | \dis(w, w') \le t \wedge \rec(w', p) \neq w] \le 2\delta$.  Thus for every value in $V|p$ this is also true.  This makes the support of $V|p$ a $(t, 2\delta)$-Shannon code~(see \defref{def:shannon-code}).  In particular, this implies that for all $w_1, w_2 \in V|p$, $\dis(w_1, w_2)\ge t$~(since $2\delta< 1/2$).  That is $V|p$ is a set with minimum distance at least $t$.  By the Singleton bound, this implies that $H_0(V | p) \le (\ell -t )|\mathbb{F}|$.  Averaging over $p$ one has that $\tilde{H}_0(V|P) \le (\ell -t +1) \log|\mathbb{F}| +1$.
%By the Markov inequality, 
%\[
%\forall w, w^*, \Pr_{p \leftarrow \sketch(w)} [ \Pr[\rec(w^*, p) = w] \le 1-2\delta ] < \delta /2.
%\]
%That is, for most $p$, $\Pr[\rec(w^*, p) = w] > 1-2\delta>1/2$.
%
%Outline of what I want to say.
%\begin{itemize}
%\item Suppose we are at some good sketch value, that is, the sketch decodes well for $w^*$ close to the true value $w$.
%\item Then for the true value that was sketched, for most neighbors $w^*$ of $w$, $\Pr[\rec(w^*, p) =w] >1/2$.  
%\item For a value $w'$ to be in the support of $W | P=p$, it must be for all neighbors of $\Pr[\rec( w', p) =w ]>1/2$.
%\item This implies that if $w_1, w_2 \in W|P = p$, then $\dis (w_1, w_2)\ge t$.
%\item Conditioned on $P=p$, $W| P=p$ is a good code for random errors.
%Consider some subset $G\subseteq supp(W| P=p)$ with $\Pr (W| P=p \in G)\ge 1/2$.  Then for all $w\in G$ the probability they recover correctly is at least $1-4\delta$~(by another application of Markov).  
%\item This, means that the size of $|G|$ is at most $n -\log |B_t|$.
%\end{itemize}
\end{proof}
We showed that for most values of $\sketch$ many values of $V$ are removed from the support.  We now show that the information contained in $Z$ is enough to nearly determine the actual sketched value.

\begin{lemma}
$\Hav(V | \sketch(V), Z) \le \Theta(1)$.
\end{lemma}
\begin{proof}
Recall that $Z$ consists of $2\ell$ coefficients and there are $(|\mathbb{F}|-1)^{\ell-1} |\mathbb{F}|^{\ell-1}$ equally likely values for $Z$.
 By described above the view of $\sketch, \rec$ is a uniform distribution $V$.  We know show there are many possible values for $Z$.  The only information seen by $\sketch$ algorithm is in the point $V=v$.  The length of this point is $|\mathbb{F}|^\ell$.  Conditioned on this information there are still many possible values for $Z$.  That is, 
 \[
 \forall v, H_0(Z | V=v) =\log \left(\frac{(|\mathbb{F}|-1)^{\ell-1} |\mathbb{F}|^{\ell-1}}{|\mathbb{F}|^\ell}\right) = \log \left( (|\mathbb{F}|-1)^{\ell-1}/|\mathbb{F}|\right).
 \]
Consider two possible $z_1, z_2$ that are possible values of $Z$.  The distributions $V| Z=z_1$ and $V | Z=z_2$ intersect at one point~(namely $v$).

This means that the $\sketch$ algorithm include points for possible $Z$ values~(other than $v$) and these values are disjoint.  The optimum strategy is to include these values uniformly from different $Z$ values.\bnote{prove this}  This means that
\begin{align*}
\Hav(V | \sketch(V), Z) &\le \tilde{H}_0(V | \sketch(V) )- \log H_0(Z | V)\\
&\le (\ell-t+1)\log |\mathbb{F}|+1 - \log (|\mathbb{F}|-1)^{\ell-1}/|\mathbb{F}|\\
&\le (\ell-t+2)\log |\mathbb{F}| - (\ell-1) \log (|\mathbb{F}|-1)+1\\
&\le (\ell-t+2)\log |\mathbb{F}| - (\ell-2) \log |\mathbb{F}|+1\\
&\le \Theta(1)
\end{align*}
 Where the last line follows for $t\ge 4$.  The fourth line follows from the third as long as $|\mathbb{F}|????$\bnote{fix this}
 \end{proof}
%\begin{lemma}
%$\Hav(V| P , Z) = \Theta(1)$
%\end{lemma}
%\begin{proof}
%Consider some fixed $p$.
%If $Z$ is universal then $\forall w_1, w_2 \in V|P=p, w_1\neq w_2$\bnote{why should $Z$ be universal for the entire range, I get for each coordinate.}
%\[
%\Pr[Z(w_1) = Z(w_2)] = \frac{1}{|Z|}.
%\]
%Then by the union bound, 
%\[
%\Pr[\exists w_1\neq w_2 \wedge Z(w_1) = Z(w_2)]\le \frac{ 2^{H_0(V|P=p)}}{|Z|}.
%\]
%\bnote{need to average across $P$}
%\bnote{I'm confused about what we mean by $Z$ being universal here.}
%\bnote{working here!}
%\end{proof}
%
%\bnote{stopped here}
\end{proof}
